---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in males. 

$H_A$: The Covid921 vaccine is effective in males. 

### Methods

The helper file used is CAT~CAT
Male and Treatments will be Cat

### Descriptive Results

### Graphical Results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=males)
barchartGC(~infected + treatment,data=males,type="percent")
```

This graph shows that the males who got the vaccine were less infected than the males who got the placebo. About 50% of both the males who got the placebo and the males who got the vaccine were okay. 

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


The chart shows that roughly 74% of the group of males which got the placebo got covid19. about 26% of the males who got the vaccine also were infected with covid19. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test and the fishers exact test are both equaling to 0.000022. This indicates that the vaccine reduces the likelihood of getting Covid 19. The odds for contracting Covid 19 increase by almost 3% by the group receiving the placebo. There is a 95% confidence interval which tells us that vaccination is effective in preventing Covid 19.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions