---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in males. 

$H_A$: The Covid921 vaccine is effective in males. 

### Methods

The helper file used is CAT~CAT.   

Male and Treatments will be Cat.  

Male and Treatments are categorical variables. 

### Descriptive Results  

### Graphical Results  

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")
```

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


The chart shows that roughly 74% of the group of males which got the placebo got covid19. about 26% of the males who got the vaccine also were infected with covid19. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test and the fishers exact test are both equaling to 0.000022. This indicates that the vaccine reduces the likelihood of getting Covid 19. The odds for contracting Covid 19 increase by almost 3% by the group receiving the placebo. There is a 95% confidence interval which tells us that vaccination is effective in preventing Covid 19.

## Females

$H_0$: The Covid921 vaccine is not effective in females. 

$H_A$: The Covid921 vaccine is effective in females. 

### Methods

The helper file used is CAT~CAT.  

Females and Treatments will be Cat.  

Females and Treatments are Categorical values.  

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")
```

This graph shows that females who got the vaccine were less infected with Covid 19 then those who were given the placebo. About 50% of the females which recieved the placebo were ok and about 51% of the females with the vaccine were ok. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that the vaccine is effective in reducing the chances of contracting covid 19 compared to the placebo.  The odds of contracting covid 19 are reduced by approximately 47% by those which received the vaccine. The vaccine lowers the risk of infection by 45% compared to the placebo. 

#### Inferential Results
 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test is equaling to 0.00001914. The Fishers Exact Test is 0.00001246 also confirmed the statistical significance of the infection rates between the placebo and the vaccination groups. The odds ratio of 1.8562 means that the odds of getting Covid within the placebo groups is 1.86 times more likely than the vaccine. There is a 95% confidence interval that suggests that the odds ratio is between 1.49 and 2.32, which tells us that the vaccine is most effective.

## LGBTQ

$H_0$: The Covid19 vaccine is not effective in the LGBTQ community.

$H_A$: The Covid19 vaccine is effective in the LGBTQ community.

### Methods

The helper files we use are Cat~Cat
LGBTQ and Treatments are Cat
LGBTQ and Treatments are categorical values

### Descriptive Results

#### Graphical Results





## Druggies


# Overall Results and Conclusions